---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r}
options(scipen=999) # turn off scientific notation
```

## Loading and preprocessing the data
```{r}
unzip("activity.zip")
dat <- read.csv("activity.csv")
head(dat)
```

## What is mean total number of steps taken per day?
```{r}
sumsteps <- tapply(dat$steps,dat$date,sum,na.rm=TRUE)
head(sumsteps)
hist(sumsteps)
mean_sumsteps <- round(mean(sumsteps),2)
median_sumsteps <- round(median(sumsteps),2)
```

The mean is ```r mean_sumsteps``` and the median is ```r median_sumsteps```

## What is the average daily activity pattern?
```{r}
avgsteps <- aggregate(dat$steps,by=list(dat$interval),mean,na.rm=TRUE)
plot(x ~ Group.1, avgsteps, xlab = "interval", ylab = "average steps", main = "Time Series plot by interval", type="l")
maxstep <- avgsteps[which.max(avgsteps$x),]
maxinterval <- maxstep$Group.1
maxvalue <- round(maxstep$x,2)
```

Maximum Interval ```r maxinterval``` and its corresponding value ```r maxvalue ```

## Imputing missing values
```{r}
sum(is.na(dat))
dat2 <- dat
```

The choosing strategy is to fill the missing value by its average value on that interval.

```{r}
dat2[is.na(dat2$steps),]$steps <- avgsteps[avgsteps$Group.1==dat2[is.na(dat2$steps),]$interval,"x"]
head(dat2)
sumsteps2 <- tapply(dat2$steps,dat2$date,sum,na.rm=TRUE)
head(sumsteps2)
hist(sumsteps2)
mean_sumsteps2 <- round(mean(sumsteps2),2)
median_sumsteps2 <- round(median(sumsteps2),2)
```

The mean is ```r mean_sumsteps2``` and the median is ```r median_sumsteps2```

The mean and median changed a bit (higher than before). The total daily number of steps increase since some missing data are filled in.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
dat2["Weekdays"] <- weekdays(as.Date(dat2$date))
dat2$Weekdays[dat2$Weekdays=="Saturday"] <- "Weekend"
dat2$Weekdays[dat2$Weekdays=="Sunday"] <- "Weekend"
dat2$Weekdays[dat2$Weekdays!="Weekend"] <- "Weekday"
dat2$Weekdays <- factor(dat2$Weekdays)

avgsteps2 <- aggregate(dat2$steps,by=list(dat2$interval,dat2$Weekdays),mean,na.rm=TRUE)
library("lattice")
xyplot(avgsteps2$x ~ avgsteps2$Group.1 | avgsteps2$Group.2,
       layout = c(1, 2), type = "l", 
       xlab = "interval", ylab = "average steps")
```
